---
title: "worm_plot_ecolgrp_by_site"
author: "Sara DeLaurentis"
date: "1/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(tidyverse)
library(glue)
library(stringr)
library(stringi)
library(magrittr)
library(dplyr)
library(wesanderson)
library(viridis)
library(ggplot2)

#install.packages("viridis")

#install.packages("wesanderson")


getwd()
```

##Temporarily set working directory to one folder back, import data from csv, summarize by site and ecol_group.
```{r }
setwd("C:/Users/sbaue/coding/R_TEMPRY/Itasca_project_19-21")

worms <- read.csv("ItascaEarthwormData_2019.csv")

```

```{r }
#For ecogroup plot (total):
worm_rep_ecogrp <- worms %>% 
  group_by(site, ecol_group) %>% 
  summarise(biomass_AFDg_TOT = sum(biomass_AFDg))
   
worm_rep_ecogrp %<>% 
  mutate(biomass_m2 = biomass_AFDg_TOT/0.080425)

worm_rep_ecogrp$inv_lvl %<>% as.character()
str(worm_rep_ecogrp)
view(worm_rep_ecogrp)
```


```{r } 
#For ecogroup plot (site mean):

worm_rep_ecogrp <- worms %>% 
  group_by(site, rep, ecol_group) %>% 
  summarise(biomass_AFDg_TOT = sum(biomass_AFDg))
   
worm_rep_ecogrp %<>% 
  mutate(biomass_m2 = biomass_AFDg_TOT/0.080425)

worm_rep_ecogrp %<>%
  group_by(site, ecol_group) %>% 
  summarise(biomass_eco_m2 = mean(biomass_m2))

#worm_rep_ecogrp$inv_lvl %<>% as.character()
#class(worm_rep_ecogrp$inv_lvl)

str(worm_rep_ecogrp)
view(worm_rep_ecogrp) 


```

#Plotting the MEAN biomass of ecological groups at each site. No error bars because that would be messy.

```{r }


wormplot.FUN <- function(data){
  Wplot <- ggplot(wormplot.df, aes(x = site, y = biomass_eco_m2, fill = ecol_group)) +
  scale_fill_manual(values = wes_palette("Darjeeling1", n = 4)) +
  geom_col(position = "stack") 
  #geom_errorbar(aes(ymin = mean-SE, ymax = mean+SE),
 #               width = .2,
  #              position = position_dodge(.9))
  Wplot <- Wplot + ggtitle("Itasca 2019 Worms", subtitle = "Mean biomass by site and ecologic group") +
    theme(plot.title = element_text(size = rel(1.8), margin = margin(b = 10)))+
    theme(axis.text.x = element_text(size = rel(1.4))) +
    theme(axis.text.y = element_text(size = rel(1.4))) 
  Wplot <- Wplot +
    labs(x= "Site", y = "EW biomass (g/m2)", size = rel(1.8)) +
    labs(fill = "Ecologic group") +
    theme(axis.title.x = element_text(size = rel(1.5), face = "bold", margin = margin(t = 8)))+
    theme(axis.title.y = element_text(size = rel(1.5), face = "bold", margin = margin(r = 8)))+
    theme(legend.text = element_text(size = rel(1.4))) +
    theme(legend.title = element_text(size = rel(1.4)))
  
 print(Wplot) #print
}

wormplot.df <- worm_rep_ecogrp
wormplot.FUN(wormplot.df) #run the function
