---
title: "worm plotting total site biomass"
author: "Sara DeLaurentis"
date: "1/15/2022"
output: html_document
purpose: create summarized worm data, take mean biomass of 3 reps at each site, calculate SD and SE, plot
related files: worm_plotting_20211027.Rmd
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(tidyverse)
library(glue)
library(stringr)
library(stringi)
library(magrittr)
library(dplyr)
library(wesanderson)
library(viridis)
library(ggplot2)
library(ggpubr)

#install.packages("viridis")

#install.packages("wesanderson")


getwd()

```
First need to summarize sums by replicate, then take the mean of the group.
Created summarized df, divided biomass totals by the worm ring area to get the biomass totals for each rep.

```{r }
#setwd("C:/Users/sbaue/coding/R_TEMPRY/Itasca_project_19-21")


#worms <- read.csv("../ItascaEarthwormData_2019.csv")
worms <- read.csv("../Itasca_worms_2021.csv")

```

## FOR 2021 WORMS ONLY ##

```{r}

worms <- worms %>% # creates column for rep.
  mutate(rep = case_when(
    grepl("W1", site_name) ~ "W1",
    grepl("W2", site_name) ~ "W2",
    grepl("W3", site_name) ~ "W3"), .after = "site"
  )

worms <- worms %>% # creates column for layer.
  mutate(layer = case_when(
    grepl("min", site_name) ~ "min",
    grepl("mus", site_name) ~ "min",
    grepl("org", site_name) ~ "org",), .after = "rep"
  )

view(worms)

#Need to select only organic horizon and mustard-extracted.

worms.org.musEX <- worms %>% 
  filter((grepl("org", site_name)) | (grepl("musEx", site_name)))

worms.org.musEX

#summarize


```

##2019 Processing starts here

```{r }

#For ecogroup plot (total):
#worm_rep_ecogrp <- worms %>% 
#  group_by(site, ecol_group) %>% 
#  summarise(biomass_AFDg_TOT = sum(biomass_AFDg))
   
#worm_rep_ecogrp %<>% 
#  mutate(biomass_m2 = biomass_AFDg_TOT/0.080425)


worm_rep_sum <-
 #worms %>% #2019
 worms.org.musEX %>% #2021
 group_by(site, rep) %>% 
 mutate(biomass_AFDg_TOT = sum(biomass.AFDG.m2)) %>% 
 ungroup() %>% 
 distinct(site, rep, .keep_all = TRUE)
  
worm_rep_sum <- worm_rep_sum %>% 
  rename("biomass_m2" = "biomass_AFDg_TOT")

worm_rep_sum$inv_lvl %<>% as.character()
#(worm_rep_sum)
view(worm_rep_sum)  

#truncate unneeded columns

keeps <- c("site", "rep", "treatment", "forest_type", "inv_lvl", "biomass_m2")
worm_rep_sum = worm_rep_sum[keeps]
view(worm_rep_sum)  

#create df to merge all desired site/rep combinations
worm.R.df <- matrix(data = NA, nrow = 24, ncol = 2)
worm.R.df[,1] <- c(rep("C2A", 3), rep("C2B", 3), rep("C5A", 3), rep("C5B", 3), rep("D2A",3), rep("D2B", 3), rep("D5A", 3), rep("D5B", 3))
#worm.R.df[,2] <- c("R1", "R2", "R3") #for 2019
worm.R.df[,2] <- c("W1", "W2", "W3") #for 2021
colnames(worm.R.df) <- c("site", "rep")

worm.R.df
#worm.R.ecogroup.df

worm_site_summary <- merge(x = worm.R.df, y = worm_rep_sum, all.x = TRUE) # merge expected values df with collected data.

na.chk <- is.na(worm_site_summary["treatment"])

#create substring column in order to rename treatment column, then delete string column
worm_site_summary$string <- substr(worm_site_summary$site, 1,2)
worm_site_summary$treatment <- worm_site_summary$string
worm_site_summary$string <- NULL

#create forest_type column from first character of treatment column
worm_site_summary$forest_type <- substr(worm_site_summary$treatment,1,1)

worm_site_summary[is.na(worm_site_summary)] <- 0 # replace NA values with 0

worm_site_summary


```

Create bar graph comparing biomass_m2 at each site. Error bars represent standard error of the mean


```{r wormplot good}

wormplot.FUN <- function(data, plot.name, title){
  Wplot <- ggplot(wormplot.stat, aes(x = site, y = mean, fill = treatment)) +
  #scale_fill_viridis(discrete = TRUE) +
  theme_classic() +
  scale_fill_grey() +
  geom_col(position = position_dodge(), colour = "black", linewidth = 1) +
  geom_errorbar(aes(ymin = mean-SE, ymax = mean+SE),
                linewidth = 1,
                width = .2,
                position = position_dodge(.9)) +
    scale_y_continuous(limits = c(0,7), expand = c(0,0))
  Wplot <- Wplot + ggtitle(title) +
    theme(plot.title = element_text(size = rel(2.5), margin = margin(b = 10)))+
    theme(axis.text.x = element_text(size = rel(2.4))) +
    theme(axis.text.y = element_text(size = rel(2.4))) 
  Wplot <- Wplot +
    labs(x= "Site", y = "EW biomass (g/m2)", size = rel(2.4), fill = "Treatment") +
    theme(axis.title.x = element_text(size = rel(2.1), face = "bold", margin = margin(t = 8)))+
    theme(axis.title.y = element_text(size = rel(2.1), face = "bold", margin = margin(r = 8)))+
    theme(legend.text = element_text(size = rel(1.7))) +
    theme(legend.title = element_text(size = rel(1.7)))
  
  png(file = glue("{plot.name}.png"), width = 600, height = 600, units = 'px') #Printing graph as png, can choose size of graph here
  print(Wplot)
  dev.off()
  return(Wplot)
  
}

#MAKING OUR INPUTS 
#1. wormplot.stat

wormplot.stat <- worm_site_summary %>% 
group_by(site) %>% 
  mutate(mean = mean(biomass_m2), SD = sd(biomass_m2)) %>% 
  ungroup() %>% 
  mutate(SE = SD/(sqrt(length(SD)))) %>% 
  distinct(site, .keep_all = TRUE)

#keep only needed columns
keeps <- c("site","treatment", "forest_type", "biomass_m2", "mean", "SD", "SE")
wormplot.stat = wormplot.stat[keeps]  

#view(wormplot.stat)

#print(wormplot.stat)

#write.csv(wormplot.stat, "wormplot.stat.2021.csv")
```

##Run the function

```{r RUN}

#2019 (Plot A)
wormplot.FUN(wormplot.stat, "mean_site_biomass_2019", "2019 Worm Biomass")
mean_site_biomass_A <- wormplot.FUN(wormplot.stat, "mean_site_biomass_2019", "2019 Worm Biomass")

print(mean_site_biomass_A)

#2021 (Plot B)
wormplot.FUN(wormplot.stat, "mean_site_biomass_2021", "2021 Worm Biomass")
mean_site_biomass_B <- wormplot.FUN(wormplot.stat, "mean_site_biomass_2021", "2021 Worm Biomass")

print(mean_site_biomass_B)

```

## Arranging plots A and B using ggarrange

```{r plot combine}

#Combining graphs using ggarrange

mean_site_biomass_combine_plot <- ggarrange(mean_site_biomass_A, #first plot (from above)
          mean_site_biomass_B + #second plot, identical to the first one except different data, has the x axis label in it because it is the center graph of the arrange
                    theme(axis.text.y = element_blank(), #removing y axis labels from this graph, I did not want more than one y axis label
                     axis.ticks.y = element_blank(),
                     axis.title.y = element_blank() ),
         # wdf_litter_plot + #third plot, identical to the first one with different data
                    #theme(axis.text.y = element_blank(), #removing y axis labels from this graph, I did not want more than one y axis label
                     #axis.ticks.y = element_blank(),
                     #axis.title.y = element_blank()),
          widths = c(0.86, 0.86), #changing widths to be equal
          common.legend = TRUE, legend = "bottom", #shared legend, legend position
          nrow = 1) #number of rows

png(file = "mean_site_biomass_combine_plot.png", width = 1400, height = 500, units = "px") #printing combined graph, can again choose size here
print(mean_site_biomass_combine_plot)


```
