#Worm plotting
#06/02/2021
#Sara DeLaurentis
#create summarized worm data

```{r }
library(tidyverse)
library(glue)
library(stringr)
library(magrittr)
library(dplyr)


getwd()
```

No NA values for data entries, but there are UNK for some of the ID values. Will need to remove those.

```{r }
worms <- read.csv("ItascaEarthwormData_2019.csv")
head(worms)
view(worms)
```

Add columns for invasion level and forest type. Relocate columns. Overwrite csv and check.

```{r }
worms <- worms %>% 
  mutate(inv_lvl = case_when(
    grepl("2", treatment) ~ "2",
    grepl("5", treatment) ~ "5"
  ), forest_type = case_when(
    grepl("C", treatment) ~ "C",
    grepl("D", treatment) ~ "D")) 

worms %<>% 
  relocate(c("forest_type", "inv_lvl"), .after = "site_name")
view(worms)
write_csv(worms, "ItascaEarthwormdata_2019.csv")

check <- read.csv("ItascaEarthwormdata_2019.csv")
view(check)
```

First need to summarize sums by replicate, then take the mean of the group.
Created summarized df, divided biomass totals by the worm ring area to get the biomass totals for each rep.

```{r }
worm_rep_sum <- 
  worms %>% 
  group_by(treatment, rep) %>% 
  summarise(biomass_AFDg_TOT = sum(biomass_AFDg))

worm_rep_sum %<>% 
  mutate(biomass_m2 = biomass_AFDg_TOT/0.080425)

view(worm_rep_sum)  
```

redo creation of invasion level and forest type columns (post-summarize)

```{r }
worm_rep_sum <- worm_rep_sum %>% 
  mutate(inv_lvl = case_when(
    grepl("2", treatment) ~ "2",
    grepl("5", treatment) ~ "5"
  ), .after = "treatment") 

worm_rep_sum <- worm_rep_sum %>% 
  mutate(forest_type = case_when(
    grepl("C", treatment) ~ "C",
    grepl("D", treatment) ~ "D"), .after = "treatment"
  )

view(worm_rep_sum)
```

Attempt to select one worm group from the original worms dataframe.

```{r }
Aporr <- worms[worms$species == "Aporrectodea species",]
view(Aporr)
```

Worm scatterplot, checking trends and relationships.

```{r }
png("worm_scatter.png", width = 1000, height = 1000)
worm_scatter <- plot(worms, cex = 2, cex.lab = 2)
print(worm_scatter)
dev.off()
?plot
```

making our own summarySE function:

```{r }
wormplot.df <- worm_rep_sum %>% group_by(treatment) %>% 
  summarise(mean = mean(biomass_m2), SD = sd(biomass_m2))

wormplot.df %<>%   #Adding in SE to the dataframe  
mutate(SE = SD/sqrt(length(SD)))
print(wormplot.df)
```


Create bar graph comparing biomass_m2 at each site. Error bars represent standard error of the mean

```{r }
wormplot.FUN <- function(data){
  Wplot <- ggplot(wormplot.df, aes(x = treatment, y = mean, fill = treatment)) +
  geom_col(position = position_dodge()) +
  geom_errorbar(aes(ymin = mean-SE, ymax = mean+SE),
                width = .2,
                position = position_dodge(.9))
  Wplot <- Wplot + ggtitle("Itasca 2019 Worms") +
    theme(axis.text.x = element_text(size = rel(1.2))) +
    theme(axis.text.y = element_text(size = rel(1.2))) +
    labs(x= "Site", y = "EW biomass / sq m")
  
 print(Wplot) #print
}

wormplot.FUN(wormplot.df) #run the function

```


anova comparing the 8 treatment groups:

```{r }
fm1 <- lm(biomass_m2 ~ treatment, data = worm_rep_sum)
anova(fm1)
summary(fm1)

view(worm_rep_sum)
```

anova comparing the 2 forest types:

```{r }
fm2 <- lm(biomass_m2 ~ forest_type, data = worm_rep_sum)
anova(fm2)
summary(fm2)
```

anova comparing the 2 invasion levels:

```{r }
fm3 <- lm(biomass_m2 ~ inv_lvl, data = worm_rep_sum)
anova(fm3)
summary(fm3)
```

plotting residuals to check for equal variance

```{r }
plot(residuals(fm1) ∼ fitted.values(fm1), main = "fm1 residuals: biomass_m2 ~ Site")
```

based on summary and residual plotting, variances are not equal.

```{r }
wormplot.df         
```

#from: http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/
#contains a potentially useful function using plyr (non-tidyverse) that I've recreated above. Maybe come back here for the CI information.
# Summarizes data.
# Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
#   data: a data frame.
#   measurevar: the name of a column that contains the variable to be summariezed
#   groupvars: a vector containing names of columns that contain grouping variables
#   na.rm: a boolean that indicates whether to ignore NA's
#   conf.interval: the percent range of the confidence interval (default is 95%)

```{r }
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(dplyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}

?ddply

```

